---
title: "Lab 6"
author: "Alissa Patterson"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank based test example

Mann-Whitney U (non-parametric alternative to an unpaired 2-sample t-test). 

H0 Null: ranks are equal (or medians)
HA Alternative: Ranks are not equal (or medians are not equal)

Are our samples so different from this null that we think they provide enough evidence that we can reject this null.

### Create samples

`set.seed()` to create "pseudorandom, but reproducible samples" 

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```


Take a quick look at the distributions using `hist()`.

```{r}
hist(gp_1)
hist(gp_2)
```

Try a t-test so we can compare the results:

```{r}
ttest <- t.test(gp_1, gp_2)

# If these data were paired, I can just add the argument `paired = "TRUE"` within the t.test ()
```

p value of .198 means that if our samples are drawn from populations with the same mean (if the null is true), there is is `r round(ttest$p.value, 2)` probability of finding sample means *at least as different* as mine by random chance.

Retain = fail to reject 

[A more comprehensive explanation of differences between groups...actually means, data spread, effect size]...there is no significant difference (t(df) = t-statistic value, p = ). In order to reject, our p would've had to be less than 5%.

### Mann Whitney U rank based test

We use the `wilcox.test()` function to perform a two-sample rank-based test.

```{r}
mwu <- wilcox.test(gp_1, gp_2)

```

H0 Null: ranks are equal (or medians)
HA Alternative: Ranks are not equal (or medians are not equal)

p value = .28, if our samples are drawn from populations with same median there is 28% of finding sample medians at least as different as ours by random chance. Cannot reject null. 

No significant difference in medians (Mann-Whitney U; U =
`r mwu$statistic`, p = `r round(mwu$p.value, 2)`)

The medians of group A (M = 22.4 g) and group B (M = 31.2 g) did not differ significantly 



